Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-03-01 paper/paper.tex
pass on technique

          symbolic x coords={method, uri, user\_agent, header, host, connection, request\_line},          

@@

        \addplot coordinates {(method,3821) (uri,10111) (user\_agent,1022) (header,4089) (host,224) (connection,29) (request\_line,11)};

@@

\Mar{----------------------------------------------------------- estou
aqui}
\subsubsection{Popularity of content modifiers.} As previously
mentioned, \percRulesWithContent\ (=26K/\numrulessuri) of Suricata
rules involve content options. Furthermore, of the 26K rules that use
content options, 13K use HTTP modifiers\Mar{@Lucas, pode dar o numero
  desconsiderando HTTP? Esta parte eh generica.}. \Luc{Posso, mas nessa função nós consideramos apenas os modificadores http.} Content modifiers
serve to indicate what part of a network packet the string mentioned
in the content option should appear. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)
Figure~\ref{fig:http-header-example} illustrates possible content
options used to match against certain strings that appear in an HTTP
message. A check (resp., cross) mark indicates a match (resp.,
unmatch).

\begin{figure}[t!]
\centering
\includegraphics[scale=0.5]{figs/http_header-example.png}
\caption{\CodeIn{http\_header} example.}
\label{fig:http-header-example}
\end{figure}

@@

\tname{} leverages the fact that these modifiers are commonly used to
define rules as to better rank the rules it synthesizes. More
precisely, it builds on the observation that strings acessible in
positions associated with content modifiers (\eg{}, GET, Host,
Connection) are more likely to produce useful rules. \tname\ uses the
frequency of content modifiers to discriminate rules that are more or
less likely to reflect actual rules. For example, historical data
indicates that content options that use the \CodeIn{http\_connection}
modifier are much less likely than content options that use the
modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. Although content
modifiers are applicable to rules of any protocol supported by
Suricata, currently, \tname\ only supports tcp and http. These are the
protocols involving ~80\% of the rules from our data set.

@@

Similarly to the the previous heuristic functions, this function uses
the frequency of content modifiers to calibrate the score of each
rule. More precisely, we compute this function with the following
formula:

@@

\noindent
, where $N$ is the number of contents options, $g(i)$ is the
contribution of content option $i$ to the score,
$\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
of a modifier $j$ to content option $i$,
$\mathit{max(modifier\_frequency)}$ refers to the modifier with
maximum popularity (for normalization), and $M_i$ is the number of
content modifiers that could access the string associated with content
$i$.

@@

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''}. The contribution of this option to the score of
the rule that uses it would be $0.5$
(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2
as the field of the HTTP message \CodeIn{Request URI}---holding the
string ``/HtmlAdaptor''---can be accessed through the modifiers
\CodeIn{http\_uri} and \CodeIn{http\_request\_line}. The relatively
high heuristic value given to this option suggests it is a plausible
option to be used in rules---the string used in the content option
(``/HtmlAdaptor'') is stored in a specific field of the HTTP message
(\CodeIn{Request URI}) and can be accessed by two content modifiers,
namely (\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the
option \CodeIn{content: ``keep-alive''}, either the modifier
\CodeIn{http\_header} or the modifier \CodeIn{http\_connection} could
be used to locate the string ``keep-alive'' stored in the field
\CodeIn{Connection} (see Figurgure~\ref{fig:http-header-example}). Consequently, for this
case, the contribution of modifiers to this content option would be
$0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
used in a rule, the aggregated contribution of this heuristic function
would be 0.35 (=(0.2+0.5)/2).
\Gui{A quinta heuristica que utilizamos considera a popularidade das opcoes. Primeiro pegamos a quantidade de vezes que uma opcao eh utilizada (sem considerar repeticoes numa unica regra). Dividimos o valor encontrado para cada opcao pelo maior valor encontrado, encontrando sempre valores de no maximo 1. Entao na regra nos somamos os valores de cada opcao, dividido pelo numero de opcoes, chegando a um numero de no maximo 1, que representa a popularidade das opcoes na regra.
\[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]}

@@

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks. Non-http attacks
    appear at the top; http attacks appear at the bottom.}
  \centering
  \begin{tabular}{llllp{10cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Source} &
    \multicolumn{1}{c}{Type} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule
    Black Nurse & \cite{pcap-attacks} & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3; \textbf{threshold}: type both, track by\_dst, count 250, seconds 1;\\    
    Ping Scan & \cite{netmap} & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & \cite{hping3} & DoS & TCP & \textbf{flags}: S,12;
    \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5;\\
    UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
    UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5; \\
    \midrule
    Apache Struts 2 & \cite{wrccdc} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\    
    ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    iisadmin access\MyComment{web-application-attack} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    WordPress Login & \cite{wrccdc} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
    .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Tomcat server snoop access & \cite{nikto} & AREC & HTTP & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  
    \bottomrule
  \end{tabular}
\end{table*}

@@

    \midrule
    Black Nurse & 1 & 4 & 9 & - \\    
    Ping Scan & 6 & 5 & 9 & 1 \\
    SYN Flood & 4 & 2 & 2 & 2 \\
    UDP Teardrop & 1 & 4 & 8 & 3 \\
    UDP Flood & - & - & - & - \\
    \midrule
    Apache Struts 2 & 2 & 65 & - & - \\
    ColdFusion & 7 & 19 & 34 & 1\\
    iisadmin access & 3 & 17 & 408 & 1 \\        
    JBoss JMX & 5 & 23 & 228 & 5 \\
    Script tag in URI & 23 & 22 & 786 & 1 \\
    WeBid LFI & 1 & 23 & 11100 & 6\\    
    Wordpress Login & 8 & 59 & - & - \\
    .htaccess access & 5 & 17 & 984 & 2\\
    .idq access & 14 & 19 & 511 & 1 \\
    /system32/ in URI & 4 & 20 & 1,913 & 1 \\
    Tomcat Snoop Access & 1 & 24 & 29830 & - \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

          symbolic x coords={uri, header, method, user\_agent, host, connection, request\_line},          

@@

        \addplot coordinates {(uri,10111) (method,3821) (user\_agent,1022) (header,4089) (host,224) (connection,29) (request\_line,11)};

@@

\subsubsection{Popularity of http content modifiers.} As previously
mentioned, \percRulesWithContent\ of \suri\ rules involve content
options. Furthermore, nearly half of the rules that use content
options, use content modifiers on http message payloads. Content
modifiers serve to indicate what part of a network packet the string
mentioned in the content option should appear. For example, the
modifier \CodeIn{http\_header} in the option \CodeIn{content:
  ``keep-alive''; http\_header} indicates that the string
``keep-alive'' should be present in the header of the
message\footnote{A keep-alive connection attribute indicates that a
  single TCP connection should remain open for multiple HTTP
  requests/responses.}.  Figure~\ref{fig:http-header-example}
illustrates possible content options used to match against certain
strings that appear in an http message. A check mark (resp., cross)
indicates a match (resp., no match).

\begin{figure}[t!]
\centering
\includegraphics[scale=0.5]{figs/http_header-example.png}
\caption{\CodeIn{http\_header} example.\Mar{Lucas, simplifique esta
    figura (reduza o tamanho da imagem). Mostre apenas os casos onde ha um match, por exemplo,
    com host=www.google.com e connection=keep-alive.}}
\label{fig:http-header-example}
\end{figure}

@@

%\tname\ uses the
%frequency of content modifiers to discriminate rules that are more or
%less likely to reflect actual rules.

\tname{} builds on our observation that these modifiers are commonly
used to define existing rules. More precisely, \tname\ builds on the
observation that the strings accessible in positions associated with
content modifiers (\eg{}, GET, Host, Connection) are more common. For
example, historical data indicates that content options that use the
modifier \CodeIn{http\_connection} are much less likely than content
options that use the modifier
\CodeIn{http\_uri}. Figure~\ref{fig:distribution-content_modifiers}
shows the histogram of content modifiers for http. (We removed the
prefix ``http'' on the modifier for space.) Although content modifiers
are applicable to rules of any protocol supported by Suricata,
currently, \tname\ only supports this features for tcp and http. But
these rules comprise $\sim$80\% of the rules of our data set.

@@

We compute the heuristic function for content modifiers with the
following formula:

@@

\noindent
, where $N$ is the number of contents options present in a rule,
$g(i)$ is the contribution of content option $i$ to the score,
$\mathit{modifier\_frequency[modifier_{ij}}]$ denotes the popularity
of a modifier $j$ to content option $i$,
$\mathit{max(modifier\_frequency)}$ refers to the modifier with
maximum popularity (for normalization), and $M_i$ is the number of
content modifiers that could access the string associated with content
$i$.

@@

To illustrate this function, consider a rule containing the option
\CodeIn{content: ``/HtmlAdaptor''}. The contribution of this option to
the score of the rule is $0.5$ (=$\frac{10,111+11}{10,111*2}$). The
value of $M_i$ for this case is 2 as the string ``/HtmlAdaptor'' could
be accessed through the modifiers \CodeIn{http\_uri} and
\CodeIn{http\_request\_line} as that string is stored on the
\CodeIn{Request URI} field of the http message\Mar{how do we know
  that?}. The relatively high heuristic value reflects the fact that
the string ``/HtmlAdaptor'', stored at \CodeIn{Request URI}, could be
accessed by two content modifiers, namely (\CodeIn{http\_uri} and
\CodeIn{http\_request\_line}).  For the option \CodeIn{content:
  ``keep-alive''}, either the modifier \CodeIn{http\_header} or the
modifier \CodeIn{http\_connection} could be used to locate the string
``keep-alive'' stored in the field \CodeIn{Connection} (see
Figure~\ref{fig:http-header-example}). Consequently, for this case,
the contribution of modifiers to this content option would be $0.2$
(=$\frac{4089+29}{10111*2}$). Assuming that both options are used in a
rule, the aggregated contribution of this heuristic function would be
0.35 (=(0.2+0.5)/2).

\subsubsection{Popularity of options.}

\Mar{--------------------------------------------------------- estou aqui}

\Gui{A quinta heuristica que utilizamos
  considera a popularidade das opcoes. Primeiro pegamos a quantidade
  de vezes que uma opcao eh utilizada (sem considerar repeticoes numa
  unica regra). Dividimos o valor encontrado para cada opcao pelo
  maior valor encontrado, encontrando sempre valores de no maximo
  1. Entao na regra nos somamos os valores de cada opcao, dividido
  pelo numero de opcoes, chegando a um numero de no maximo 1, que
  representa a popularidade das opcoes na regra.
\[\frac{\sum_{i=1}^{N}\frac{\mathit{option\_frequency[option_i]}}{\mathit{max(option\_frequency)}}}{N}\]}

@@

\begin{table*}[t!]
  \small
  \caption{\label{table:attacks}List of attacks. Non-http attacks
    appear at the top; http attacks appear at the bottom.}
  \centering
  \begin{tabular}{llllp{10cm}}
    \toprule
    \multicolumn{1}{c}{Name} &
    \multicolumn{1}{c}{Source} &
    \multicolumn{1}{c}{Type} &
    \multicolumn{1}{c}{Protocol} &
    \multicolumn{1}{c}{Groud Truth (relevant/checked options)} \\
    \midrule
    Black Nurse & \cite{pcap-attacks} & DoS & ICMP & \textbf{icode}:3; \textbf{itype}:3; \textbf{threshold}: type both, track by\_dst, count 250, seconds 1;\\    
    Ping Scan & \cite{netmap} & AREC & ICMP & \textbf{dsize}:0; \textbf{itype}:8; \\
    SYN Flood & \cite{hping3} & DoS & TCP & \textbf{flags}: S,12;
    \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5;\\
    UDP Teardrop & \Fix{WireShark} & DoS & UDP & \textbf{fragbits}:M; \textbf{id}:124; \\
    %%UDP Flood & MazeBolt & DoS & UDP & \textbf{fragbits}:M; \textbf{threshold}: type both, track by\_dst, count 5000, seconds
    5; \\
    \midrule
    %%Apache Struts 2 & \cite{wrccdc} & RCE & HTTP & \textbf{content}:"POST"; http\_method; \textbf{content}:"java.lang.ProcessBuilder"; nocase; http\_client\_body; fast\_pattern; \textbf{content}:"/struts2-rest-showcase/orders/3"; http\_uri; \\    
    ColdFusion \MyComment{PESC} & \cite{nikto} & PESC & HTTP \MyComment{& Privilege Escalation} & \textbf{content}:"GET"; http\_method; nocase; \textbf{content}:"/CFIDE/administrator"; http\_uri; nocase; \\
    iisadmin access\MyComment{web-application-attack} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"/iisadmin"; nocase; http\_uri; \\
    JBoss JMX\MyComment{ Console Beanshell Exploit} & \cite{nikto} & RCE & HTTP \MyComment{& Remote Code Execution} & \textbf{content}:"/HtmlAdaptor"; nocase; http\_uri; \textbf{content}:"action=inspect"; nocase; http\_uri; \textbf{content}:"bean"; nocase; http\_uri; \textbf{content}:"name="; http\_uri; \\
    Script tag in URI\MyComment{Possible Cross Site Scripting Attempt} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:"</script>"; nocase; http\_uri; \\
    WeBid LFI \MyComment{Local File Inclusion} & \cite{nikto} & LFI & HTTP & \textbf{content}:"GET "; depth:4; \textbf{content}:"/cron.php?"; http\_uri; nocase; \textbf{content}:"include\_path="; http\_uri; nocase; \textbf{content}:"../"; \\
    %%WordPress Login & \cite{wrccdc} & PVE & HTTP & \textbf{content}:"log="; http\_client\_body; \textbf{content}:"\&pwd="; http\_client\_body; \textbf{content}:"\&wp-submit="; http\_client\_body; \\
    .htaccess access\MyComment{attempted-recon} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:".htaccess"; nocase; http\_uri; \\
    .idq access\MyComment{EXPLOIT ISAPI} & \cite{nikto} & RCE & HTTP \MyComment{& Adicionar} & \textbf{content}:".idq"; nocase; http\_uri; \\
    /system32/ in URI\MyComment{Possible Protected Directory Access Attempt} & \cite{nikto} & AREC & HTTP \MyComment{& Adicionar} & \textbf{content}:"/system32/"; nocase; http\_uri; \\
    Tomcat server snoop access & \cite{nikto} & AREC & HTTP & \textbf{content}:"/jsp/snp/"; http\_uri; \textbf{content}:".snp"; http\_uri;\\  
    \bottomrule
  \end{tabular}
\end{table*}

@@

    \midrule
    Black Nurse & 1 & 4 & 9 & 2 \\    
    Ping Scan & 6 & 5 & 9 & 2 \\
    SYN Flood & 4 & 2 & 2 & 2 \\
    UDP Teardrop & 1 & 4 & 8 & 4 \\
    %%UDP Flood & - & - & - & - \\
    \midrule
    %%Apache Struts 2 & 2 & 65 & - & - \\
    ColdFusion & 7 & 19 & 34 & 3\\
    iisadmin access & 3 & 17 & 408 & 1 \\        
    JBoss JMX & 5 & 23 & 228 & 45 \\
    Script tag in URI & 23 & 22 & 786 & 1 \\
    WeBid LFI & 1 & 23 & 11100 & 10\\    
    %%Wordpress Login & 8 & 59 & - & - \\
    .htaccess access & 5 & 17 & 984 & 2\\
    .idq access & 14 & 19 & 511 & 2 \\
    /system32/ in URI & 4 & 20 & 1,913 & 4 \\
    Basebuilder LFI & 1 & 22 & 18133 & 34 \\
    Tomcat Snoop Access & 1 & 24 & 29830 & 32 \\
    \bottomrule
  \end{tabular}
\end{table}
